``Order is Half of Life'' -- Establishing Totally Ordered HTN Planning Problems is Half the Job (to Getting them Solved Efficiently)

### 2022-12-23 paper/01--introduction.tex

\section{Introduction}
% \input{01--intro.tex}
Hierarchical Task Network (HTN) planning is a hierarchical approach to planning. As defined by \cite{HTNSurvey},  %  and \cite{IntroGhallab},
tasks in HTN planning are either primitive, corresponding to an action that can be taken, or compound. HTN problems have a set of methods that specify how one might achieve a given compound task, by decomposing it into a set of sub-tasks. A compound task may even decompose into itself, either directly via a method, or indirectly via a sequence of method applications. If decomposition leads to a sequence of primitive tasks executable from the initial state, then this sequence of actions is a solution to the problem, also known as a \enquote{plan}.

@@

% Certain kinds of problems might be naturally more suited to being modelled as a partially ordered problem, for example, the actions \enquote{deliver package 1 to city A} and \enquote{deliver package 2 to city B} -- these are essentially unrelated goals in a real transport scenario, and so modelling the problem to require that one task be completed before the other would be unnecessarily limiting the possible solution space. For some problems, such over-specification may remove all valid solutions.

@@

Another benefit of transforming a POHTN problem to a TOHTN problem is that the additional structure to the problem could allow us to deploy specialised algorithms and heuristics for the totally ordered case. % Also, heuristic design is comparably easy for total-order problems due to the missing interaction between tasks.
The drawback to this approach is that, due to the greater expressivity of POHTN planning, there may exist POHTN problems that cannot be solved when converted to a TOHTN problem.
Fortunately, not every POHTN problem is guaranteed to be undecidable, and so could still be transformed while preserving at least one solution.
In this paper we present and investigate an algorithm for converting POHTN problems to TOHTN problems. We prove that when certain criteria are met, it guarantees that at least one solution will be preserved.
Also, we obtain a new class of decidable problems, namely those that satisfy the above mentioned criterion. Finally, we show that, even when these criteria are not met, very few problems are rendered unsolvable by the transformation, and that it greatly reduces solving time for problems, with gains being bigger for more difficult problems.

>>>>

\section{Introduction}

Hierarchical Task Network (HTN) planning is a hierarchical approach to planning. As defined by \cite{HTNSurvey},
tasks in HTN planning are either primitive, corresponding to an action that can be taken, or compound. HTN problems have a set of methods that specify how one might achieve a given compound task, by decomposing it into a set of sub-tasks. A compound task may even decompose into itself, either directly via a method, or indirectly via a sequence of method applications. If decomposition leads to a sequence of primitive tasks executable from the initial state, then this sequence of actions is a solution to the problem, also known as a \emph{plan}.

@@

@@

Another benefit of transforming a POHTN problem to a TOHTN problem is that the additional structure to the problem could allow us to deploy specialised algorithms and heuristics for the totally ordered case.
The drawback to this approach is that, due to the greater expressivity of POHTN planning, there may exist POHTN problems that cannot be solved when converted to a TOHTN problem.
Fortunately, not every POHTN problem is guaranteed to be undecidable, and so could still be transformed while preserving at least one solution.
In this paper we present and investigate an algorithm for converting POHTN problems to TOHTN problems. We prove that when certain criteria are met, it guarantees that at least one solution will be preserved.
Also, we obtain a new class of decidable problems, namely those that satisfy the above mentioned criterion. Finally, we show that, even when these criteria are not met, very few problems are rendered unsolvable by the transformation, and that it greatly reduces solving time for problems, with gains being bigger for more difficult problems.
